\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-15 paper.qmd
Updated Figure 2

<!--langevitour with true model for S-curve-->
::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"}

![](figures/scurve/sc_true_1.png){width="150" fig-align="center"}

![](figures/scurve/sc_true_2.png){width="150" fig-align="center"}

![](figures/scurve/sc_true_3.png){width="150" fig-align="center"}

Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>). 
:::

@@

true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")

>>>>

```{r}
#| label: s-curve-true-model-proj1

true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
wireframe_true_model <- read_rds("data/s_curve/scurve_true_model_wireframe.rds")

projection <- cbind(
    c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
    c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))

projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection

projected_true_df <- projected_true |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

model_df <- dplyr::left_join(
  wireframe_true_model, 
  projected_true_df, 
  by = c("from" = "ID"))

names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")

scurve_proj1_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#969696")

```{r}
#| label: s-curve-true-model-proj2

projection <- cbind(
    c(0.03685,-0.02239,0.08047,0.22415,-0.10295,0.04529,0.10381),
    c(-0.16908,0.11691,0.04886,0.10189,-0.00203,-0.09763,-0.13207))

projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection

projected_true_df <- projected_true |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

model_df <- dplyr::left_join(
  wireframe_true_model, 
  projected_true_df, 
  by = c("from" = "ID"))

names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")

scurve_proj2_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#969696")

```{r}
#| label: s-curve-true-model-proj3

projection <- cbind(
    c(0.11484,0.06997,-0.13638,0.01096,0.01533,0.16257,-0.13615),
    c(0.07805,0.01652,0.17174,-0.07343,0.16177,0.11112,0.04729))

projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection

projected_true_df <- projected_true |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

model_df <- dplyr::left_join(
  wireframe_true_model, 
  projected_true_df, 
  by = c("from" = "ID"))

names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")

scurve_proj3_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#969696")

```{r}
#| label: fig-scurve-proj
#| fig-cap: "Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."

scurve_proj1_true_model + scurve_proj2_true_model + 
  scurve_proj3_true_model + plot_layout(ncol = 3)
```

<!-- <!--langevitour with true model for S-curve--> -->
<!-- ::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"} -->

<!-- ![](figures/scurve/sc_true_1.png){width="150" fig-align="center"} -->

<!-- ![](figures/scurve/sc_true_2.png){width="150" fig-align="center"} -->

<!-- ![](figures/scurve/sc_true_3.png){width="150" fig-align="center"} -->

<!-- Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>).  -->
<!-- ::: -->

@@

### 2024-08-15 paper.tex
Updated Figure 2

\begin{minipage}{0.33\linewidth}
\begin{center}
\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_1.png}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\begin{center}
\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_2.png}
\end{center}
\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\begin{center}
\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_3.png}
\end{center}
\end{minipage}%

\caption{\label{fig-scurve-true-sc}Three views of the true model (grey
points and lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for
the S-curve data (purple points). The data is spread along the S shape,
and does not vary much from this curve. (The \textbf{langevitour}
software is used to view the data with a tour, and the full video is
available at \url{https://youtu.be/I5GL23vLiw0}).}

@@

@@

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-24-1.pdf}
\end{center}

@@

Edges define the neighbourhood structure, in order to provide a smooth
\gD{} representation of the fitted model.
Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
that was used to generate the S-curve example data. The ideal is that
the representation of the fitted model, at least for this example where
we know the true model, should look similar to this.

@@

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-57-1.pdf}
\end{center}

>>>>

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-scurve-proj-1.pdf}

\caption{\label{fig-scurve-proj}Three views of the true model (grey
points and lines) in \(2   ext{-}D\) projections from \(7    ext{-}D\),
for the S-curve data (purple points). The data is spread along the S
shape, and does not vary much from this curve. (The \textbf{langevitour}
software is used to view the data with a tour, and the full video is
available at \url{https://youtu.be/I5GL23vLiw0}).}

@@

--\textgreater{}

@@

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-28-1.pdf}
\end{center}

@@

Edges define the neighbourhood structure, in order to provide a smooth
\gD{} representation of the fitted model. \textbf{?@fig-scurve-true-sc}
shows a wire frame of the true model that was used to generate the
S-curve example data. The ideal is that the representation of the fitted
model, at least for this example where we know the true model, should
look similar to this.

@@

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-61-1.pdf}
\end{center}
